### 2016-08-16 rebuttal.tex

\begin{verbatim}
   >>> (exp(100)+1).evalf() - exp(100).evalf()
   0
   >>> ((exp(100)+1) - exp(100)).evalf()
   1.00000000000000
   >>> (exp(100)+1) - exp(100)
   1
\end{verbatim}

or two different ways to compute the 100th Fibonacci number:

\begin{verbatim}
   >>> phi = (1+sqrt(5))/2
   >>> psi = (1-sqrt(5))/2
   >>> ((phi**100-psi**100)/sqrt(5)).evalf()-fibonacci(100)
   65536.0000000000
   >>> ((phi**100-psi**100)/sqrt(5) - fibonacci(100)).evalf()
   0.e-104
\end{verbatim}

Obviously, like your own example, these are problematic because a part of the computation is relying on Python's builtin floats. Please include a comment on whether symbolic computation (i.e., applying \texttt{evalf()} on the whole expression) always avoids these errors or not, possibly with an example when it doesn't resolve this problem.
\begin{solution}
First your example, unfortunately, doesn't work due to automatic
simplification (see last input line --- \texttt{exp(100)} cancel).
Second example is good, but our looks to be more simple.

Applying \texttt{evalf()} has nothing related to \emph{symbolic} computation,
that method is for numerical evaluation, that should be clear from
the sentence below ``The evalf method converts a constant symbolic expression
to a Float\dots''.  Numerical evaluation of the whole expression should
always solve cancellation problem, as it was stated in the sentence
``Applying the evalf method to the whole expression solves this problem.''
\end{solution}

\question The footnote from line 477 should be moved as a sentence in its own right to the introduction, with other technical specifications.
\begin{solution}
Mpmath version now mentioned in the introduction.
\end{solution}

### 2016-08-16 rebuttal.tex

\question I suggest removing the \verb|>>>| before each line. It looks strange in a notebook.
\begin{solution}
We have done this.
\end{solution}
\end{questions}
\subsection{Comments for the author}
Sympy is a superb package and I am happy to see that there is now a paper that describes its current state. Here are a few comments that you may wish to consider before publication.

\begin{questions}
\question Line 105
It is generally frowned upon to import all symbols from a Python module in this manner.
I understand why you are doing it in this paper, it makes subsequent sympy commands less verbose.
It might, however, encourage bad practice. It may also lead to a poor user experience for newbies.

For example, say I had the following code

\begin{verbatim}
𝔬mνmpyimportsin,arraytest=array(123)sin(test)
\end{verbatim}
[sic]

@@

I decide that I want to use sympy for something and, following your example, do

\begin{verbatim}
𝔬mνmpyimportsin,array𝔬msympyimport⋅test=array(123)sin(test)x,y,z=symbols(′xyz′)
\end{verbatim}
[sic]

@@

This reads badly. Perhaps the following might be better?

This is due, in part, to the fact that the same language, Python, is used both for the internal implementation and the external usage by users.

\begin{solution}
We have changed the sentence as suggested.
\end{solution}
\question Line 221:
the phrase `Expression tree' is cited but this is not the first time you've used it. Perhaps cite earlier? Line 129 perhaps?

\begin{solution}
The footnote here was in reference to the SymPy classes. We have moved the
footnote later in the sentence, to ``\texttt{Basic}'', as suggested
by reviewer 3 (point \ref{rev3point16}).
\end{solution}
\question Page 8: Footnote 5
The line reads

@@

\question The word ``software'' in line 91 is ambiguous; ``library'' or ``package'' would make a better choice.
\begin{solution}

\end{solution}

\question Lines 91--96: ``we discuss/look at/etc'' is the preferred form, instead of ``section discusses/looks at/etc''.
\begin{solution}
  % TODO: We had decided not to use first person. We should perhaps email the editor
  % to ask about this.
\end{solution}

\question The paragraph in lines 103--105 should be moved to the introduction,
and the footnote from line 104 should be added to that paragraph as a
full-blown sentence, expanded by all the relevant technical information
(Python version, OS, \ldots). Given that the end of life for Python 2 is
2020., a comment on whether all the presented examples work in both Python 2
and 3 should be included as well. Further, emphasise that wildcard imports,
``\texttt{import *}'', should almost never be used in programs (see PEP 8, the
item ``wildcard imports''). The same goes for the \texttt{import} mentioned in lines 484--487.
\begin{solution}

\question Line 119 should be removed, as it is basically a copy of the previous line.
\begin{solution}
Line 119 is the output of line 118, so now showing it would
be incorrect.

@@

\question Line 119 should be removed, as it is basically a copy of the previous line.
\begin{solution}
Line 119 is the output of line 118, so now showing it would
be incorrect.

The aim here was simply to show that the input expression remains unevaluated.
We have added a sentence before the example to note this.
This particular example was chosen because it shows the basic syntax for
addition, subtraction, multiplication, division, and exponentiation.
\end{solution}

\question In line 121 the word ``stored'' should be replaced by ``used as keys''.
\begin{solution}

\question What do the authors mean by ``thereby permitting features such as caching'' in line 122? Caching can be done for mutable types as well, just not through hashing.
\begin{solution}

\question There is no need to repeat ``(CAS)'' in line 124, as it was already given in line 64.
\begin{solution}

\question In the same line, the word ``represents'' should be replaced by ``stores''.
\begin{solution}
  % I don't personally agree with this, but it's a minor point. --Aaron
\end{solution}

\question In line 184, ``symbols are'' should be replaced by ``\texttt{t} is'' (the general rule is already given in line 179).
\begin{solution}

\question The code in line 210 should be made into its own line (like a displaymath formula), for typesetting reasons and better readability.
\begin{solution}

@@

\question The supplement should have an introduction, explaining what is being covered in it in general and in each section.
\begin{solution}

\end{solution}

\question Since the Guntz algorithm is covered in depth, it would be good to
include how an interested reader can see SymPy's steps of computation:

\begin{verbatim}
     >>> import os
     >>> os.environ['SYMPY_DEBUG'] = 'True'
     >>> from sympy import *
     >>> x = symbols('x')
     >>> limit(sin(x)/x, x, 0)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(0, 1, None, None)] with symbol _w
     DEBUG: returned None
     DEBUG: parsing of expression [(_w, 1, None, None)] with symbol _w
     DEBUG: returned ([], [(_w, 1, None, None)], 1, False)
     DEBUG: parsing of expression [(1, 1, None, None)] with symbol _w
     DEBUG: returned None
     limitinf(x*sin(1/x), x) = 1
     +-mrv_leadterm(_p*sin(1/_p), _p) = (1, 0)
     | +-mrv(_p*sin(1/_p), _p) = ({_p: _Dummy_14}, {}, _Dummy_14*sin(1/_Dummy_14))
     | | +-mrv(_p, _p) = ({_p: _Dummy_14}, {}, _Dummy_14)
     | | +-mrv(sin(1/_p), _p) = ({_p: _Dummy_15}, {}, sin(1/_Dummy_15))
     | |   +-mrv(1/_p, _p) = ({_p: _Dummy_15}, {}, 1/_Dummy_15)
     | |     +-mrv(_p, _p) = ({_p: _Dummy_15}, {}, _Dummy_15)
     | +-rewrite(_Dummy_14*sin(1/_Dummy_14), {exp(_p): _Dummy_14}, {}, _p, _w)
             = (sin(_w)/_w, -_p)
     | | +-sign(_p, _p) = 1
     | | +-limitinf(1, _p) = 1
     | +-calculate_series(sin(_w)/_w, _w) = 1
     |   +-limitinf(_w, _w) = oo
     |   | +-mrv_leadterm(_w, _w) = (1, -1)
     |   | | +-mrv(_w, _w) = ({_w: _Dummy_18}, {}, _Dummy_18)
     |   | | +-rewrite(_Dummy_18, {exp(_w): _Dummy_18}, {}, _w, _w) = (1/_w, -_w)
     |   | | | +-sign(_w,_w)=1
     |   | | | +-limitinf(1, _w) = 1
     |   | | +-calculate_series(1/_w, _w) = 1/_w
     |   | +-sign(-1, _w) = -1
     |   | +-sign(1, _w) = 1
     |   +-limitinf(_w, _w) = oo
     |   +-limitinf(_w, _w) = oo
     |   +-limitinf(_w, _w) = oo
     +-sign(0, _p) = 0
     +-limitinf(1, _p) = 1
\end{verbatim}

### 2016-08-16 introduction.tex
Address question 26 of the reviewer 3
Change "printers" to "display formatters" in the introduction
The concept of printer hasn't been introduced yet at this point. This
addresses a comment by reviewer 3.

Unlike many CASs, SymPy does not invent its own programming language. Python
itself is used both for the internal implementation and end user
interaction. By using the operator overloading functionality of Python, SymPy follows the embedded domain specific language paradigm proposed by Hudak~\cite{dsl-little-languages}.  The exclusive usage of a single programming language makes it easier
for people already familiar with that language to use or develop SymPy.
Simultaneously, it enables developers to focus on mathematics, rather than
language design.  SymPy officially supports Python 2.6, 2.7 and 3.2--3.5.

@@

As a library, SymPy does not have a built-in graphical user interface (GUI).
However, SymPy exposes a rich interactive display system, and supports
registering display formatters with Jupyter~\cite{kluyver2016jupyter} frontends,
including the Notebook and Qt Console, which will render SymPy expressions
using MathJax~\cite{cervone2012mathjax} or \LaTeX{}.

The remainder of this paper discusses key components of the SymPy library.
Section~\ref{sec:architecture} discusses the architecture of SymPy.
Section~\ref{sec:features} enumerates the features of SymPy and takes a closer
look at some of the important ones. The section~\ref{sec:numerics}
looks at the numerical features of SymPy and its dependency library, mpmath.
Section~\ref{sec:domain_specific} looks at the domain specific physics
submodules for performing symbolic and numerical calculations in classical mechanics
and quantum mechanics. Conclusions and future directions for SymPy are given
in section~\ref{sec:conclusion}.  All examples in this paper use SymPy version 1.0.

The following statement imports all SymPy functions into the global Python
namespace.\footnote{\texttt{import *} has been used here to aid the
  readability of the paper, but is best to avoid such wildcard import
  statements in production code, as they make it unclear which names are
  present in the namespace. Furthermore, imported names could clash with
  already existing imports from another package. For example, SymPy, the
  standard Python \texttt{math} library, and NumPy all define the \texttt{exp}
  function, but only the SymPy one will work with SymPy symbolic expressions.}
From here on, all examples in this paper assume that this statement has been
executed:

\begin{verbatim}
>>> from sympy import *
\end{verbatim}

>>>>

Unlike many CASs, SymPy does not invent its own programming language. Python
itself is used both for the internal implementation and end user
interaction. By using the operator overloading functionality of Python, SymPy follows the embedded domain specific language paradigm proposed by Hudak~\cite{dsl-little-languages}.  The exclusive usage of a single programming language makes it easier
for people already familiar with that language to use or develop SymPy.
Simultaneously, it enables developers to focus on mathematics, rather than
language design.

@@

As a library, SymPy does not have a built-in graphical user interface (GUI).
However, SymPy exposes a rich interactive display system, and supports
registering display formatters with Jupyter~\cite{perez2007ipython} frontends,
including the Notebook and Qt Console, which will render SymPy expressions
using MathJax~\cite{cervone2012mathjax} or \LaTeX{}.

The remainder of this paper discusses key components of the SymPy software.
Section~\ref{sec:architecture} discusses the architecture of SymPy.
Section~\ref{sec:features} enumerates the features of SymPy and takes a closer
look at some of the important ones. The section~\ref{sec:numerics}
looks at the numerical features of SymPy and its dependency library, mpmath.
Section~\ref{sec:domain_specific} looks at the domain specific physics
submodules for performing symbolic and numerical calculations in classical mechanics
and quantum mechanics. Conclusions and future directions for SymPy are given
in section~\ref{sec:conclusion}.
